---
title: "HARVARD EXTENSION SCHOOL"
subtitle: "EXT CSCI E-106 Model Data Class Group Project Template"
author:
- Author One
- Author Two
- Author Three
- Author Four
- Author Five
- Author Six

tags: [logistic, neuronal networks, etc..]
abstract: |
  This is the location for your abstract.

  It must consist of two paragraphs.
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry: margin=1.3cm
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---
\newpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
HouseSales<-read.csv("KC_House_Sales.csv")
```
\newpage
## House Sales in King County, USA data to be used in the Final Project

| Variable| Description |
| :-------:| :------- |
| id| **Unique ID for each home sold (it is not a predictor)**    |
| date| *Date of the home sale*    |
| price| *Price of each home sold*    |
| bedrooms| *Number of bedrooms*    |
| bathrooms| *Number of bathrooms, where ".5" accounts for a bathroom with a toilet but no shower*    |
| sqft_living| *Square footage of the apartment interior living space*    |
| sqft_lot| *Square footage of the land space*    |
| floors| *Number of floors*    |
| waterfront| *A dummy variable for whether the apartment was overlooking the waterfront or not*    |
| view| *An index from 0 to 4 of how good the view of the property was*    |
| condition| *An index from 1 to 5 on the condition of the apartment,*    |
| grade| *An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 has a high-quality level of construction and design.*    |
| sqft_above| *The square footage of the interior housing space that is above ground level*    | 
| sqft_basement| *The square footage of the interior housing space that is below ground level*    |
| yr_built| *The year the house was initially built*    |
| yr_renovated| *The year of the house’s last renovation*    |
| zipcode| *What zipcode area the house is in*    |
| lat| *Latitude*    |
| long| *Longitude*    |
| sqft_living15| *The square footage of interior housing living space for the nearest 15 neighbors*    |
| sqft_lot15| *The square footage of the land lots of the nearest 15 neighbors*    |
\newpage
## Instructions:
0.  Join a team with your fellow students with appropriate size (Four Students total)
1.  Load and Review the dataset named "KC_House_Sales'csv
2.	Create the train data set which contains 70% of the data and use set.seed (1023). The remaining 30% will be your test data set.
3.	Investigate the data and combine the level of categorical variables if needed and drop variables as needed. For example, you can drop id, Latitude, Longitude, etc.
4.	Build a regression model to predict price. 
5.	Create scatter plots and a correlation matrix for the train data set. Interpret the possible relationship between the response.
6.	Build the best multiple linear models by using the stepwise selection method. Compare the performance of the best two linear models. 
7.	Make sure that model assumption(s) are checked for the final model. Apply remedy measures (transformation, etc.) that helps satisfy the assumptions. 
8.	Investigate unequal variances and multicollinearity. If necessary, apply remedial methods (WLS, Ridge, Elastic Net, Lasso, etc.). 
9.	Build an alternative model based on one of the following approaches to predict price: regression tree, NN, or SVM.  Check the applicable model assumptions. Explore using a logistic regression. 
10.	Use the test data set to assess the model performances from above.
11.	Based on the performances on both train and test data sets, determine your primary (champion) model and the other model which would be your benchmark model.
12.	Create a model development document that describes the model following this template, input the name of the authors, Harvard IDs, the name of the Group, all of your code and calculations, etc..:

## Due Date: December 18th, 2023 at 11:59 pm EST

**Notes**
**No typographical errors, grammar mistakes, or misspelled words, use English language**
**All tables need to be numbered and describe their content in the body of the document**
**All figures/graphs need to be numbered and describe their content**
**All results must be accurate and clearly explained for a casual reviewer to fully understand their purpose and impact**
**Submit both the RMD markdown file and PDF with the sections with appropriate explanations. A more formal document in Word can be used in place of the pdf file but must include all appropriate explanations.**

Executive Summary

This section will describe the model usage, your conclusions and any regulatory and internal requirements. In a real world scneario, this section is for senior management who do not need to know the details. They need to know high level (the purpose of the model, limitations of the model and any issues).


\newpage
## I. Introduction (5 points)

*This section needs to introduce the reader to the problem to be resolved, the purpose, and the scope of the statistical testing applied. What you are doing with your prediction? What is the purpose of the model? What methods were trained on the data, how large is the test sample, and how did you build the model?*


\newpage
## I. Description of the data and quality (15 points)

*Here you need to review your data, the statistical test applied to understand the predictors and the response and how are they correlated. Extensive graph analysis is recommended. Is the data continuous, or categorical, do any transformation needed? Do you need dummies? *


\newpage
## III. Model Development Process (15 points)

*Build a regression model to predict price.  And of course,  create the train data set which contains 70% of the data and use set.seed (1023). The remaining 30% will be your test data set. Investigate the data and combine the level of categorical variables if needed and drop variables. For example, you can drop id, Latitude, Longitude, etc. *

\newpage
## IV. Model Performance Testing (15 points)

*Use the test data set to assess the model performances. Here, build the best multiple linear models by using the stepwise both ways selection method. Compare the performance of the best two linear models. Make sure that model assumption(s) are checked for the final linear model. Apply remedy measures (transformation, etc.) that helps satisfy the assumptions. In particular you must deeply investigate unequal variances and multicollinearity. If necessary, apply remedial methods (WLS, Ridge, Elastic Net, Lasso, etc.). *


\newpage
## V. Challenger Models (15 points)
*Build an alternative model based on one of the following approaches to predict price: regression tree, NN, or SVM. Explore using a logistic regression. Check the applicable model assumptions. Apply in-sample and out-of-sample testing, backtesting and review the comparative goodness of fit of the candidate models. Describe step by step your procedure to get to the best model and why you believe it is fit for purpose.*

We first built a basic regression tree model as a baseline non-parametric challenger model; results were not particularly strong (r-squared at ~0.6, which is lower than the r-squared of the linear regression model). We then tuned hyperparameters for the decision tree using random search (using RMSE and MAE as the metrics to optimise) and backtested the model using a cross-validation resampling approach. This only improved performance very slightly (small decline in RMSE and improvement in r-squared to 0.62). Ultimately we chose this tuned model as our challenger model, as it is relatively simple to build and implement, and fairly explainable (e.g. the splits tend to be easy to understand). More complex models e.g. deep neural networks would likely perform better, but would likely consume too many computational resources for our purposes.

```{r}
# Load libraries
library(caret)
library(rpart)
library(nnet)
library(knitr)
library(dplyr)
library(stringr)
library(readr)
library(kableExtra)
library(mlr)
library(rpart.plot)

# Model building

# Build basic regression tree
tree_model <- rpart(price ~ ., data = train)

# Visualize the decision tree
par(mfrow=c(1,2))  
rpart.plot(tree_model, shadow.col = "gray", nn = TRUE)

# Make predictions using basic regression tree
tree_predictions <- predict(tree_model, newdata = test)
tree_rmse <- sqrt(mean((tree_predictions - test$price)^2))
tree_mae <- mean(abs(tree_predictions - test$price))
tree_mse <- mean((tree_predictions - test$price)^2)

#View basic regression tree results 
cat("Regression Tree RMSE:", tree_rmse, "\n")
cat("Regression Tree MAE:", tree_mae, "\n")
cat("Regression Tree MSE:", tree_mse, "\n")

#Tune regression tree hyperparameters to improve performance 

# Define the hyperparameter space for random search
param_space <- makeParamSet(
  makeNumericParam("cp", lower = 0.01, upper = 0.1),
  makeIntegerParam("minsplit", lower = 5, upper = 10),
  makeIntegerParam("minbucket", lower = 5, upper = 10),
  makeIntegerParam("maxdepth", lower = 5, upper = 15),
  makeIntegerParam("maxcompete", lower = 0, upper = 1)
)

# Create a learner for regression with rpart
learner <- makeLearner("regr.rpart", predict.type = "response")

# Define the backtesting strategy (in this case, 10-fold cross-validation)
resampling <- makeResampleDesc("CV", iters = 10)

# Define the random search control object
ctrl <- makeTuneControlRandom(maxit = 10) 

# Perform random search for hyperparameter tuning
random_search <- tuneParams(
  learner = learner,
  task = makeRegrTask(data = train, target = "price"),
  resampling = resampling,
  par.set = param_space,
  control = ctrl,
  measures = list(rmse, mae)  # Specify evaluation metrics
)

# Get the best hyperparameters
best_hyperparameters <- random_search$x

# Print the best-tuned hyperparameters
print(best_hyperparameters)

# Train the final model using the best hyperparameters 
tree_model_tuned <- rpart(
  price ~ ., 
  data = train,
  control = rpart.control(
    cp = best_hyperparameters$cp,
    minsplit = best_hyperparameters$minsplit,
    minbucket = best_hyperparameters$minbucket,
    maxdepth = best_hyperparameters$maxdepth,
    maxcompete = best_hyperparameters$maxcompete
  )
)

# Make predictions using the final tuned model
predictions <- predict(tree_model_tuned, newdata = test)

# Calculate residuals
residuals <- test$price - predictions
# Create diagnostic plots using 
par(mfrow = c(2, 2))  

# Scatter plot of residuals vs. predictions 
ggplot(data = data.frame(Predictions = predictions, Residuals = residuals)) +
  geom_point(aes(x = Predictions, y = Residuals)) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs. Predictions",
       x = "Predictions",
       y = "Residuals") +
  theme_minimal()

# Histogram of residuals 
ggplot(data = data.frame(Residuals = residuals)) +
  geom_histogram(aes(x = Residuals), bins = 30, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Residuals",
       x = "Residuals",
       y = "Frequency") +
  theme_minimal()

# Normal Q-Q plot of residuals 
qqplot_data <- data.frame(Standardized_Residuals = scale(residuals))
ggplot(qqplot_data, aes(sample = Standardized_Residuals)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  labs(title = "Normal Q-Q Plot of Residuals") +
  theme_minimal()

# Residuals vs. Index plot 
ggplot(data = data.frame(Residuals = residuals, Index = seq_along(residuals))) +
  geom_point(aes(x = Index, y = Residuals)) +
  labs(title = "Residuals vs. Index",
       x = "Index",
       y = "Residuals") +
  theme_minimal()

# Paired t-test for differences between observed and actual values
t_test <- t.test(test$price, tree_predictions, paired = TRUE)
cat("\nPaired t-test for Differences:\n")
cat("t statistic:", t_test$statistic, "\n")
cat("p-value:", t_test$p.value, "\n")
if (t_test$p.value < 0.05) {
  cat("There is a significant difference between observed and predicted values (reject null hypothesis).\n")
} else {
  cat("There is no significant difference between observed and predicted values (fail to reject null hypothesis).\n")
}

# Compare model performance
compare_model_performance <- function(models, model_names, train, test) {
  rmse_values <- c()
  rsquared_values <- c()
  mape_values <- c()
  
  for (i in 1:length(models)) {
    model <- models[[i]]
    predictions <- predict(model, newdata = test)
    rmse <- sqrt(mean((test$price - predictions)^2))
    rsquared <- cor(test$price, predictions)^2
    mape <- mean(abs((test$price - predictions) / test$price)) * 100
    
    rmse_values <- c(rmse_values, rmse)
    rsquared_values <- c(rsquared_values, rsquared)
    mape_values <- c(mape_values, mape)
  }
  
  return(list(rmse_values, rsquared_values, mape_values))
}

# Call the function to compare model performance
models <- list(tree_model_tuned, tree_model)  
model_names <- c("Tuned regression tree model", "Basic regression tree model")  
performance_metrics <- compare_model_performance(models, model_names, train, test)

# Create a data frame to hold the results
model_results <- data.frame(Model = model_names, 
                             RMSE = performance_metrics[[1]], 
                             R_squared = performance_metrics[[2]], 
                             MAPE = performance_metrics[[3]])

# Create a table to compare model results
kable(model_results, format = "markdown")
```

\newpage
## VI. Model Limitation and Assumptions (15 points)

*Based on the performances on both train and test data sets, determine your primary (champion) model and the other model which would be your benchmark model. Validate your models using the test sample. Do the residuals look normal? Does it matter given your technique? How is the prediction performance using Pseudo R^2, SSE, RMSE?  Benchmark the model against alternatives. How good is the relative fit? Are there any serious violations of the model assumptions? Has the model had issues or limitations that the user must know? (Which assumptions are needed to support the Champion model?)* 


\newpage
## VII. Ongoing Model Monitoring Plan (5 points)

*How would you picture the model needing to be monitored, which quantitative thresholds and triggers would you set to decide when the model needs to be replaced? What are the assumptions that the model must comply with for its continuous use?*


\newpage
## VIII. Conclusion (5 points)

*Summarize your results here. What is the best model for the data and why?*

## Bibliography (7 points)

*Please include all references, articles and papers in this section.*

## Appendix (3 points)

*Please add any additional supporting graphs, plots and data analysis.*


